\documentclass{article}
\nonstopmode

% Pacakges
%
\usepackage[margin=0.75in]{geometry}
\usepackage{fancyvrb}
\usepackage{multicol}

% Common Declarations
%
% Make sure you change the author in common.tex
%
\input{../common.tex}

% Local Declarations
%
\title{\course, Lab \# 1}
\author{\term}
\date{}

% The document
%
\begin{document}
\maketitle

\begin{multicols}{2}
  
\begin{description}
\item[Goal:] The purpose of this lab is write some simple code in {\em
  Scheme}, and to get familiar with the hardware and software that we
  will be using all quarter, the lab room, your TA, and the
  sumbmission procedure with GitLab.

\item[Due:] Your program, named {\tt lab1.rkt}, must be committed and pushed to
  GitLab before midnight, Friday, January 15.

\item[Unit tests:] At a minimum, your program must pass the unit tests found in
  the file {\tt lab1-test.rkt}.  It will report any error; there should be no
  output.

\item[Program:] Write a {\sc Scheme} procedure called {\tt make-pi} to compute
  $\pi$ using the (slowly converging) series:
  \[
    {\pi} = 4 - \frac43 + \frac45 - \frac47 + \frac49 - \ldots
  \]
  The procedure takes one parameter, which will be the accuracy we need.  We can
  stop whenever the next factor we would add is smaller than this accuracy.  For
  example:
\begin{Verbatim}[frame=single]
(make-pi 0.1) =>  3.09162380666784
(make-pi 0.001) =>  3.1410926536210413
(make-pi 0.0000001) => 3.141592603589817
\end{Verbatim}
  This series converges {\em very} slowly.  If you are curious, instrument the
  procedure so that it also prints the number of iterations it took to get the
  accuracy.

  Make sure that your program {\em returns} the value of $\pi$ computed, and
  doesn't just print it.  For example, this should not give an error:
  \verb|(+ (make-pi 0.1) (make-pi 0.1))|


  To get the loop done, define a recursive procedure with (at least) three
  parameters that behave like this (I've truncated the decimals):\\
  \begin{tabular}{rrr}
    Numerator & Denominator & Sum   \\\hline
    4.0       & 1.0         & 0.0   \\
    -4.0      & 3.0         & 4.0   \\
    4.0       & 5.0         & 2.666 \\
    -4.0      & 7.0         & 3.466 \\
    4.0       & 9.0         & 2.895 \\
    -4.0      & 11.0        & 3.339 \\
    4.0       & 13.0        & 2.976 \\
    -4.0      & 15.0        & 3.283 \\
    4.0       & 17.0        & 3.017 \\
    -4.0      & 19.0        & 3.252 \\
    4.0       & 21.0        & 3.041 \\
  \end{tabular}

  Make sure your program starts its loop with floating point numbers, {\em e.g.}
  4.0, 1.0, {\em etc.}  If you start with exact integers, Scheme will try to
  keep exact rational numbers through all of those computations and it will be
  substantially slower.  Also, your answers will look like this:
\begin{Verbatim}[frame=single]
(make-pi 0.1) => 
    516197940314096/166966608033225
\end{Verbatim}

  Make sure to follow the general rules about functional programming: write lots
  of functions, no assignment statements, no {\tt for} or {\tt while} loops, use
  {\tt cond} instead of {\tt if}.

  Name your program {\tt lab1.rkt} and include a comment block like the one
  shown with your own name and W-number.  Also, hopefully, with the correct
  code, although this one will pass the unit tests!

  \VerbatimInput[frame=single,label=lab1.rkt]{lab1.rkt}

\end{description}
\end{multicols}

\end{document}

\end
