\documentclass[letterpaper,12pt]{article}
\nonstopmode

% Packages
%
\usepackage[margin=1in]{geometry}
\usepackage{fancyvrb,amsmath}
\usepackage{multicol}
\usepackage{xcolor}

% Common Declarations
%
% Make sure you change the author in common.tex
%
\input{../common.tex}
\fvset{frame=single,fontsize=\small}
\setlength{\columnsep}{2em}

% Local Declarations
% 
\title{CSCI 301, Lab \# 5}
\author{\term}

% The document
%
\begin{document}

\maketitle
  \paragraph{Goal:} This is the fourth in a series of labs that will build an
  interpreter for Scheme.  In this lab we will add the {\tt lambda} special
  form.

  \paragraph{Unit tests:} At a minimum, your program must pass the unit tests
  found in the file {\tt lab6-test.rkt}.  Place this file in the same folder as
  your program, and run it; there should be no output.  Include your unit tests
  in your submission.

  \paragraph{Lambda creates a closure:} If you consider a simple application of
  a {\tt lambda} form, it is a lot like {\tt let}, that you did last week:
\begin{Verbatim}
((lambda (x y) (+ x y)) (+ 2 2) (+ 3 3)) 
   <=> (let ((x (+ 2 2)) (y (+ 3 3))) (+ x y))
\end{Verbatim}
  In each case, a list of arguments, \texttt{((+ 2 2) (+ 3 3))} is evaluated
  giving the list of values \texttt{(4 6)}, then
  the list of variables, \texttt{(x y)}, is bound to the list
  of values \texttt{(4 6)}, and this is used to extend the environment,
  and then the expression, \texttt{(+ x y)}, is evaluated in the extended
  environment, resulting in a value of 10 for the whole expression.

  If this were the only use of \texttt{lambda}, then it would just be a
  different syntax for \texttt{let}.

  However, the \texttt{lambda} is actually more powerful.  A \texttt{lambda} can
  be created in one environment, and applied in another.  Consider the following
  code:
\begin{Verbatim}[frame=single]
(let ((a 10) (f (let ((a 20)) (lambda (x) (+ a x)))))
     (f 30))
\end{Verbatim}
  The body of the function \texttt{f} is \texttt{(+ a x)}.  The function was
  called with argument 30, so \texttt{x} should be bound to 30.  But what is
  \texttt{a}?  Is the answer $10+30$ or $20+30$?

  You can check it out with Racket, but hopefully you can see that it
  \emph{should} be $20 + 30$.  When the \texttt{lambda} form was
  \emph{evaluated}, \texttt{a} was bound to 20.  When the resulting function,
  bound to \texttt{f}, was \emph{applied}, \texttt{a} was 10.

  The solution to this is that a \texttt{lambda} form creates a special data
  structure called a \emph{closure}.  The closure is what is bound to \texttt{f},
  in the above example.  A closure consists of three things:
  \begin{enumerate}
  \item A list of symbols (the variables to be bound when it is called).
  \item An expression to be evaluated.
  \item A stored environment, the environment in which the closure was
    \emph{created}.
  \end{enumerate}
  The first two are easy to understand.  For example, in the \texttt{lambda}
  form: \texttt{(lambda (x y) (+ x y))}, the list of symbols is \texttt{(x y)}
  and the expression is \texttt{(+ x y)}.  The third thing is the environment
  that was in effect when the closure was created.  In other words, when the
  \texttt{lambda} form was \emph{evaluated}.  A closure, when created, stores
  this third thing, the current environment, along with the list of variables
  and the body.

  Because this is lisp, we will represent closures simply as lists containing
  the three items.  Since this is going to be a data type, let's go ahead and
  define all the functions that will use closures, a creator, a predicate, and
  three accessors:
\begin{Verbatim}[frame=single,label=Closure Data Type]
(define closure (lambda (vars body env) (list 'closure vars body env)))
(define closure? (lambda (clos) (and (pair? clos) (eq? (car clos) 'closure))))
(define closure-vars cadr)
(define closure-body caddr)
(define closure-env cadddr)  
\end{Verbatim}
  Make sure in your code you {\em only} use these procedures to handle closures.
  Respect the interface!  (If you're interested, you can look up \textbf{Racket}
  {\em structures}, which could have been used instead of lists.)  Note that
  since closures are just lists, we don't need to write any special printing
  procedures to look at them.  They print nicely as lists already.

  So, \texttt{lambda} is the new special form for this assignment.  Evaluating a
  lambda form simply creates a closure out of the arguments, the body, and the
  current environment.  For example, if the environment \texttt{e1} is created
  like this:
\begin{Verbatim}[commandchars=\\\{\}]
(define e1{\color{blue} '((x 5) (y 8) (z 10))} )
\end{Verbatim}
  then the following closures would look like this:
\begin{Verbatim}[frame=single,commandchars=\\\{\}]
(evaluate  '(lambda (x) (+ x y)) e1)
=> (closure (x) (+ x y) 
     {\color{blue}((x 5) (y 8) (z 10))})
(evaluate  '(lambda (a b c) 
              (cons a (list b c))) e1)
=>  (closure (a b c) (cons a (list b c)) 
       {\color{blue}((x 5) (y 8) (z 10)})
(evaluate '(let ((x 10)) 
     (lambda (foo) (+ foo foo)) e1)
=>  (closure (foo) (+ foo foo) ((x 10) 
        {\color{blue}(x 5) (y 8) (z 10))})
\end{Verbatim}
  Note that the closure remembers the environment in which it was created, even
  if that environment was a special environment, for example, like the one
  created by a \texttt{let} form.

  \paragraph{Applying lambda to some arguments:}
  Now that we know how to evaluate a \texttt{lambda} form, we need to know how
  to {\em apply} a lambda form to some arguments.  For example, consider the
  form
\begin{Verbatim}[frame=single]
((lambda (x y) (+ x y)) 10 20)
\end{Verbatim}
  This is {\em not} a \texttt{lambda} form.  It is not even a special form!  It
  is the {\em application} of the \texttt{lambda} form \texttt{(lambda (x y) (+
    x y))} to the arguments, \texttt{(10 20)}.  We will evaluate this just like
  any other function application!

  If we follow our rules for evaluating things that are not special forms, we
  would evaluate each of the items in the list (the \texttt{lambda} form, the 10
  and the 20), getting this list
  (printing the closure as a list):\\
  \begin{tabular}{|rcccl|}\hline
    \texttt{(}&\texttt{(lambda (x y) (+ x y))} & 10 & 20 & \texttt{)}\\
              &  \ensuremath{\Downarrow} & \ensuremath{\Downarrow} & \ensuremath{\Downarrow} &\\
    \texttt{(} & \texttt{(closure (x y) (+ x y) {\color{blue}((x 5) (y 8) (z
                 10) ...)}) } & 10 &  20& \texttt{)}\\\hline
  \end{tabular}\\
  And now we have to apply the closure to the list of arguments.  Up to
  now, we've only been applying Racket built-in functions to their
  arguments, using the Racket \texttt{apply} function.  However, Racket
  doesn't understand our closures.  We have to figure out how to apply
  them ourselves.

  So, we will add a new \texttt{apply-function} procedure to our interpreter.
  When evaluating a normal (non-special-form) list, we evaluate each item in the
  list, and then call \texttt{apply-function} with just two arguments: the
  \texttt{car} and the \texttt{cdr} of our evaluated list of items, the function
  and its arguments.

  \texttt{apply-function} will look at its first argument.  If it is a \texttt{
    procedure?}, then it calls the Racket built-in \texttt{apply} function.  If
  it is a \texttt{closure?}, it calls \texttt{apply-closure}.  Otherwise it
  should throw an error, reporting an unknown function type.

  \texttt{apply-closure} takes two arguments, a closure and a list of values.
  The closure has three components: the variables, the body, and a saved
  environment.  This procedure extends the {\em saved} environment by appending
  the variables and their values to the front (just like \texttt{let} did), and
  then evaluates the body of the closure in this new, extended environment.

  Note that \texttt{let} is not the only form that introduces a local
  environment, now \texttt{lambda} does, too.  Consider the following code:
\begin{Verbatim}[frame=single]
(let ((f (lambda (a) (lambda (b) (+ a b)))))
  (let ((g (f 10))
        (h (f 20)))
    (list (g 100) (h 100))))  
\end{Verbatim}
  Here, the result of applying \texttt{f} to 10 creates a closure in an
  environment in which \texttt{a} is bound to 10.  When we apply \texttt{f} to
  20, we create a closure in an environment in which \texttt{a} is bound to 20.
  What should be the result of the call?  Does your interpreter get that result?

  Note also that we can't define recursive functions directly using \texttt{
    let}.  What would happen if we tried this?
\begin{Verbatim}[frame=single]
  (let ((f
           (lambda (n)
              (if (= n 0)
                  1
                  (* n (f (- n 1)))))))
      (f 5))
\end{Verbatim}
  Try this in the \textbf{Racket} interpreter, and also in your interpreter.  In
  our next lab, we will define the \texttt{letrec} special form to remedy this.

  However, we don't really need to wait for that to define recursive functions.
  We just have to be a little more clever.  For example, what would the
  following do in \textbf{Racket}?  What does it do in your interpreter?
\begin{Verbatim}[frame=single]
(let ((f
       (lambda (f n)
         (if (= n 0)
             1
             (* n (f f (- n 1)))))))
  (f f 5))
\end{Verbatim}
\end{document}


